---
title: "Lab 3: Spatial Basics"
subtitle: <h4 style="font-style:normal">GEOG-364 - Spatial Analysis</h4>
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: true
    theme: flatly
    section_numbering: false
---



```{r, include=FALSE, echo=FALSE,results='hide',warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)

# THE FORMAT YOU LIKE IS LAB 3 INSIDE 2022

# invisible data read
library(tidyverse)
library(sp)
library(sf)
library(readxl)
library(skimr)
library(ggplot2)
library(tmap)
library(viridis)
library(kableExtra)
library(plotly)
library(readxl)
library(osmdata)



```

<br>

## Welcome to Lab 3!

The aim of this lab is to start looking at tables of data and to make your first maps.

By the end of this week's lab, you will be able to:

-   

    1.  Make a markdown template

-   

    2.  Read in data from an excel file

-   

    3.  Conduct exploratory data analysis

-   

    4.  Wrangle a table of data

-   

    5.  Make initial plots and maps

The lab is worth 110 points and there is a rubric at the end of this page.

```{=html}
<p class="comment"><strong>THIS IS A TWO WEEK LAB. You have two full lab sessions, then, assignment 3 is due the week afterwards.</strong></p>
```
<br><br>


## A: Set up - DON'T SKIP

<br>

#### Step A1. Files & Projects

<details>
<summary>POSIT CLOUD people, expand for set up</summary>
<br>

**Step i:**<br>Go to <https://posit.cloud/content/> and make a new project for Lab 3 <br><br>

**Step ii:**<br>Run this code IN THE CONSOLE to install the packages you need. <br>

 -  This is going to take 5-10 minutes, so let it run and carry on. <br><br>
 

```{r, eval=FALSE}

install.packages("readxl")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")

install.packages("remotes")

remotes::install_github(repo = "r-spatial/sf", 
                        ref = "93a25fd8e2f5c6af7c080f92141cb2b765a04a84")

install.packages("terra")
install.packages("tmap")
install.packages("elevatr")
install.packages("nasapower")
install.packages("osmdata")
install.packages("tigris")
install.packages("rmdformats")
install.packages("plotly")
install.packages("ggstatsplot")



# If you get a weird error it might be that the quote marks have messed up. 
# Replace all of them and ask Dr G for a shortcut.

```
<br><br>

**Step iii:   **<br>Go to the Lab Canvas page and download these 3 files.  Upload each one into your project folder.<br>*Forgotten how? See Lab 2 Set-Up.* <br>

   1. Your PERSONAL UFO dataset (look for the csv file with your email ID). <br>
   2. Your PERSONAL Spatial State-level dataset. <br>
   3. The lab report template .RmD file.<br>
   
     
<br>
</details>
 
<br>
<details>
<summary>R-DESKTOP people, expand for set up</summary>

<br>

**Step i:**<br>Create a new project for Lab 3.<br>*Forgotten how? See Lab 2 Set-Up.*<br><br>

**Step ii:**<br>Go to the Lab Canvas page and download these 3 files.<br> Put each one in your Lab 3 project folder. <br>

   1. Your PERSONAL UFO dataset (look for the csv file with your email ID). <br>
   2. Your PERSONAL Spatial State-level dataset. <br>
   3. The lab report template .RmD file.<br><br>
   
**Step iii:**<br>Open the Lab report template and try to knit. Install any packages you are missing (either via clicking the little yellow bar or by clicking Packages/Install)  <br><br>

</details>
<br><br>

#### Step A2. Change the theme


<details>
<summary>Instructions</summary>
<br>

This is more for us when we're grading!  It helps us to not go insane when we're reading 60 reports.<br><br>

**Step i:**  Change the AUTHOR line to your personal E-mail ID <br><br>
 
**Step ii:**  Go here and choose a theme out of downcute, robobook, material, readthedown or html_clean: <https://github.com/juba/rmdformats?tab=readme-ov-file#formats-gallery>.<br> *DO NOT CHOOSE html_docco - it doesn't have a table of contents*  <br><br>
 
**Step iii:**<br>Change the theme line on the template YAML code to a theme of your choice. See the example YAML code below <br>
*Note, downcute chaos is different - see below*<br><br>

 1. Example YAML code if you want `robobook`, `material`, `readthedown` or `html_clean`. You can also change the highlight option to change how code chunks look. - see the rmdformats website.

```{r,eval=FALSE}

---
title: "Lab 3"
author: "ADD YOUR EMAIL ID"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    self_contained: true
    highlight: kate
---

```

<br>

 2. Example YAML code if you want `downcute` or `downcute chaos`. For white standard downcute, remove the downcute_theme line.

```{r,eval=FALSE}

---
title: "Lab 3"
author: "ADD YOUR EMAIL ID"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    downcute_theme: "chaos"
    default_style: "dark"
---
      
```

<br><br>
</details>
<br><br>

#### Step A3. DO NOT KNIT

There is code missing in your reports for you to complete and it will not work.


<br><br>

## B: UFO Analysis - Week 1

You have each been given some (genuine) about UFOs for a specific US State.  Because you all have unique data, you will all get different results and so you can work together.

<br>

#### Step B2. Learn about the datasets


<details>
<summary>Instructions</summary>
<br>


**Step i:** 
First, learn about the UFO data (you will write about it in Step 3). Look at these websites and BRIEFLY note down to yourself,, how was the data collected? Why? e.g. what were the people who collected the data trying to achieve?  Also note down any other relevant information you can see. E.g. is data stored as points or polygons etc.  <br>

 - The raw data was collected from here: <https://nuforc.org/>  <br>
 
  - The actual data was obtained from here <https://github.com/planetsig/ufo-reports>  <br>
 

<br>
**Step ii:** 
You are also going to be using some data from the US government American Community Survey (you will write about it in Step 3). Look at these resources and BRIEFLY note down to yourself, how was the data collected? Why? e.g. what were the people who collected the data trying to achieve?  Also note down any other relevant information you can see. E.g. is data stored as points or polygons etc.  <br>

 - You can read about it here: <https://www.socialexplorer.com/help/faq/knowledge-base/american-community-survey-basics>  <br>
 
  - Or there are a LOAD of youtube videos and resources out there  <br>
 

<br><br>
</details>
<br><br>

#### Step B3. Follow the lab report instructions, up to and including "UFO Non Spatial Summary"
<br>
<details>
<summary>Tutorials to come. Instructions also in Lab Report Template</summary>
<br>
</details>
<br><br>

#### Step B4. Now knit. Check it works


<br><br>

#### END OF WEEK 1 Work
My apologies. I know this is a short lab and that it pushes the hard thinking to next week.   Getting the new materials up which work with the new R map projection has been a real pain, but I'm getting there!

<br>

```{=html}
<p class="comment"><strong>REALLY focus on writing up your data sources well - and getting on top of this week's homework and discussion. <br><br> Also, go to google maps and get to know your assigned US State!  Where are its cities/its roads/anything else relevant?.  Turn on the satellite/terrain layers and look at the landscape.  This will all help you understand the spatial patterns you might see in the UFO data</strong></p>
```


<br><br>


#### Do I need to submit this week?

You don't need to submit anything this week.

HOWEVER, I will give one bonus point (within the lab) to anyone who submits their completed .Rmd and .html up until this point in your Lab 3 canvas assignment. 

 - You are getting the bonus point for not leaving it all until the last minute.  I don't think I have the capacity to provide feedback unless you ask questions as a submission comment
  - You will resubmit with the full report next week




