---
title: "Lab 5 Tutorial Template"
author: "hlg5155"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    highlight: tango
---


```{r, message=FALSE, warning=FALSE}
# Add all your libraries here
library(sf)
```

# Instructions

You can remove these at the end

Also remember that YOU WON'T NEED EVERY CODE CHUNK.  For example you don't need any of the stuff on rearranging the data because I have already done that for you.   I have marked the sections where you might find the answers in the hints.

Your tutorial should include the following,  

 - An explanation of what autocorrelation is (negative/zero/positive) and (first-second order) - Hint B1

 - Why traditional statistical tests become challenging when working with data that exhibits spatial autocorrelation. - Hint B2

 - A brief description and map of a single variable of your choice, commenting what the data is showing and what processes might have caused any pattern you see (population density? climate?) - Hint Section D.

 - A global moran's scatterplot and test for that variable, both standard and monte carlo - explaining what is going on in each case (Hint sections E, F and G, ignoring the renaming part in F)

 - A LISA analysis for that variable, explaining how the LISA map links back to your moran's scatterplot (Hint section H)


REMEMBER YOU HAVE ALREADY DONE MOST OF THIS WORK!  (e.g. you can copy/paste ANYTHING you wrote in the lab so far) - and remember the readings on canvas or the lecture notes (feel free to include screenshots).  The coding should also be easy because you are copy/pasting your code from this example into your new report and just changing the table/variable name.  e.g. instead of PennACS_sf$total.pop you might have Richmond_ACS$med.age




# Reading in the data

Remember to put Richmond_ACS.gpkg into your project folder - and then this code should run. You can carry on from here.

```{r}
Richmond_ACS <- st_read("Richmond_ACS.gpkg")
```


```{r}
Richmond_ACS
```


# Your tutorial..

## What is autocorrelation